Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-22 paper/sections/proof-reconstruction/conjunct.tex
[ paper ] clausify added.
last changes.

\begin{mainth}
  \label{thm:conjunct}
  If $Γ ⊢ φ$ and $ψ  : \Prop$ then $Γ ⊢ \fconjunct~φ~ψ$, where,
  \begin{equation*}
  % \label{eq:rank-definition}
  \begin{aligned}
  &\hspace{.495mm}\fconjunct : \Prop \to \Prop \to \Prop\\
  &\begin{array}{llll}
  \fconjunct~φ~ψ &=
        \begin{cases}
            ψ, &\text{ if }φ ≡ ψ\\
            ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂\text{ and }\fconjunct(φ₁, ψ) ≡ ψ  \\
            ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂\text{ and }\fconjunct(φ₂, ψ) ≡ ψ  \\
            φ, &\text{ otherwise.}
          \end{cases}
  \end{array}
  \end{aligned}
  \end{equation*}
\end{mainth}

\begin{proof}\hspace{2cm}\\
Case $φ ≡ ψ$. $Γ ⊢ \fconjunct~φ~ψ$ normalizes to $Γ ⊢ ψ$.
Then, we get the desire conclusion by applying the $\fsubst$ theorem.\\[2mm]
Case $φ ≡ φ₁ ∧ φ₂$. If we can get $ψ ≡ \fconjunct(φ_{i}, ψ)$ for some $i = 1,\ 2$, then,

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ₁ ∧ φ₂$}
  \RightLabel{∧-proj$_{i}$}
  \UnaryInfC{$Γ ⊢ φ_{i}$}
  \RightLabel{by hyp.}
  \UnaryInfC{$Γ ⊢ \fconjunct(φ_{i}, ψ)$}
  \AxiomC{$ψ ≡ \fconjunct(φ_{i}, ψ)$}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
Otherwise, the last case follows trivially.
\end{proof}

>>>>

\begin{mainth}
  \label{thm:conjunct}
  If $Γ ⊢ φ$ and $ψ  : \Prop$ then $Γ ⊢ \fconjunct~φ~ψ$, where,
  \begin{equation*}
  % \label{eq:rank-definition}
  \begin{aligned}
  &\hspace{.495mm}\fconjunct : \Prop \to \Prop \to \Prop\\
  &\begin{array}{llll}
  \fconjunct~φ~ψ &=
        \begin{cases}
            ψ, &\text{ if }φ ≡ ψ\\
            ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂\text{ and }\fconjunct~φ₁~ψ ≡ ψ  \\
            ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂\text{ and }\fconjunct~φ₂~ψ ≡ ψ  \\
            φ, &\text{ otherwise.}
          \end{cases}
  \end{array}
  \end{aligned}
  \end{equation*}
\end{mainth}

\begin{proof}\hspace{2cm}\\
Case $φ ≡ ψ$. $Γ ⊢ \fconjunct~φ~ψ$ normalizes to $Γ ⊢ ψ$.
Then, we get the desire conclusion by applying the $\fsubst$ lemma.\\[2mm]
Case $φ ≡ φ₁ ∧ φ₂$. If we can get $ψ ≡ \fconjunct~φ_{i}~ψ$ for some $i = 1,\ 2$, then,

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ φ₁ ∧ φ₂$}
  \RightLabel{∧-proj$_{i}$}
  \UnaryInfC{$Γ ⊢ φ_{i}$}
  \RightLabel{by hyp.}
  \UnaryInfC{$Γ ⊢ \fconjunct~φ_{i}~ψ$}
  \AxiomC{$ψ ≡ \fconjunct~φ_{i}~ψ$}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}
Otherwise, the last case follows trivially.
\end{proof}

### 2017-10-22 paper/sections/proof-reconstruction/resolve.tex
last changes.

\begin{mainlemma}
\label{lem:lem-rassoc}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rassoc}_{⬓}~φ$.
\end{mainlemma}

\begin{equation*}
  \begin{split}
  \label{eq:build-or}
  &\rm{build}_{∨} : \Prop → \Prop → \Prop\\
  &\rm{build}_{∨}(φ, ψ) =%\\
  % &\hspace{3mm}
  \begin{cases}
  ψ, &\text{ if } φ ≡ ψ\\
  ψ, &\text{ if } ψ ≡ ψ₁ ∨ ψ₂\text{, } \rm{build}_{∨}(φ, ψ_{i}) ≡ ψ_{i}\text{ for some }i=1, 2\\
  φ, &\text{ otherwise.}
  \end{cases}
  \end{split}
\end{equation*}

@@

\begin{mainlemma}[lem-build$_{\boldsymbol{∨}}$]
\label{lem:lem-build-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{build}_{∨}~φ~ψ$.
\end{mainlemma}

@@

\begin{mainlemma}[lem-factor]
\label{lem:lem-factor}
  $Γ ⊢ φ$ then $Γ ⊢ \ffactor(φ)$.
\end{mainlemma}

@@

\begin{mainlemma}
\label{lem:lem-sbuild-or}
If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{sbuild}_{∨}(φ, ψ)$.
\end{mainlemma}

@@

\begin{mainth}[thm-reorder$_{\boldsymbol{∨}}$]
  \label{thm-reorder-or}
  If $Γ ⊢ φ$ and $ψ: \Prop$ then $Γ ⊢ \freorder_{∨}~φ~ψ~$, where,
  \begin{align*}
    \label{eq:reorder-or}
    \begin{split}
    &\freorder_{∨} : \Prop \to \Prop \to \Prop\\
    &\freorder_{∨}~φ~ψ~= \rm{sbuild}_{∨}~(rassoc_{∨}~φ)~ψ.
    \end{split}
  \end{align*}
\end{mainth}

\begin{proof}
Use Lemma~\ref{lem:lem-rassoc} and Lemma~\ref{lem:lem-sbuild-or}.
\end{proof}

@@

\begin{mainlemma}[lem-reorder$_{\boldsymbol{∧}}$]
  \label{lem:lem-reorder-and}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧}(φ, ψ)$.
\end{mainlemma}

@@

\begin{mainlemma}[lem-conjunct$_{\boldsymbol{∨}}$]
  \label{lem:lem-conjunct-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fconjunct_{∨}(φ, ψ)$.
\end{mainlemma}

\begin{definition}[reorder$_{\boldsymbol{∧∨}}$]
 \begin{align*}
      \begin{split}
      &\freorder : \Prop \to \Prop \to \Prop\\
      &\freorder_{∧∨}(φ, ψ)=
        \begin{cases}
          ψ, &\text{ if } φ≡ψ\\
          ψ, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂,\ \freorder_{∧∨}(φ,ψ₁) ≡ ψ₁\\
             &\text{and }\freorder_{∧∨}(φ, ψ₂) ≡ ψ₂\\
          φ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂\\
          \fconjunct_{∨}(φ, ψ) &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
  \end{definition}

\begin{mainth}[thm-reorder$_{\boldsymbol{∧∨}}$]
  \label{thm:reorder-and-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧∨}(φ, ψ)$.
\end{mainth}

@@

To emulate this rule, we perform a sequence of rearrangements inside the
formulas involved to match with the pattern expected by the resolution theorem
described above. We begin by defining a convenient version of
the binary resolution in Def.~\ref{def:rsol}.

\begin{definition}[rsol]
\label{def:rsol}
\begin{align*}
      \begin{split}
        &\rm{rsol} : \Prop \to \Prop\\
        &\rm{rsol}(φ) =
        \begin{cases}
          φ₂, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁, φ₂ ≡ φ₄\\
          φ₂ ∨ φ₄, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁\\
          φ, &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{mainlemma}[lem-rsol]
  \label{lem:lem-rsol}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rsol}(φ)$.
\end{mainlemma}

@@

\begin{definition}[resolve]
\label{def:resolve}
 \begin{align*}
 \begin{split}
    &\fresolve : \Prop \to \Prop \to \Prop \to \Prop \to \Prop\\
    &\fresolve(φ₁, φ₂, l, ψ) =
      \rm{rsol}(\freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨} (φ₂, ¬ l ∨ ψ)).
  \end{split}
  \end{align*}
\end{definition}

\begin{mainth}[thm-resolve]
  \label{thm:resolve}
  If $Γ ⊢ φ₁$, $Γ ⊢ φ₂$, a literal $l$, and $ψ : \Prop$ then
  $Γ ⊢ \fresolve(φ₁, φ₂, l, ψ)$.
\end{mainth}

\Needspace{3\baselineskip}
\begin{proof}
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₁$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
    \vskip 1.5mm
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₂$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
  \vskip 1.5mm
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ)$}
    \AxiomC{$Γ ⊢ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{∧-intro}
    \BinaryInfC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{lem-rsol.}
    \UnaryInfC{$Γ ⊢ \rm{rsol}(\freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨}(φ₂, ¬ l ∨ ψ))$}
  \end{bprooftree}
  \end{equation*}
\end{proof}

\begin{figure}
\label{fig:resolve-example}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ ¬ r ∨ p ∨ q$}
  \AxiomC{$Γ ⊢ p ∨ q ∨ r $}
  \RightLabel{thm-resolve}
  \BinaryInfC{$Γ ⊢ \fresolve(¬ r ∨ p ∨ q, p ∨ q ∨ r, r, p ∨ q)$}
  \UnaryInfC{$Γ ⊢ p ∨ q$}
  \end{bprooftree}
\end{equation*}
\caption{Example of a resolution in \Metis using
Thm.~\ref{thm:resolve}.}
%Application of Thm.~\ref{thm:resolve} for the \TSTP excerpt Fig.~\ref{fig:resolve-tstp-example}.}
\end{figure}

>>>>

\begin{mainlemma}
\label{lem:rassoc}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rassoc}_{⬓}~φ$.
\end{mainlemma}

\begin{equation*}
 \label{eq:build-or}
  \begin{aligned}
  &\hspace{.495mm}\rm{build}_{∨} : \Prop → \Prop → \Prop\\
  &\begin{array}{llll}
    \rm{build}_{∨}&ψ~ψ &= ψ\\
    \rm{build}_{∨}&(φ~(ψ₁ ∨ ψ₂)&=ψ &\text{ if }\rm{build}_{∨}~φ~ψ_{1} ≡ ψ_{1}\\
    \rm{build}_{∨}&(φ~(ψ₁ ∨ ψ₂)&=ψ &\text{ if }\rm{build}_{∨}~φ~ψ_{2} ≡ ψ_{2}\\
    \rm{build}_{∨}&φ~ψ&=φ&\text{ otherwise.}
   \end{array}
  \end{aligned}
\end{equation*}

\begin{equation}
  \label{eq:ndisj-or}
  \begin{aligned}
  &\hspace{.495mm}\rm{ndisj}_{∨} : \NProp \to \NProp\\
  &\begin{array}{lll}
    \rm{ndisj}_{∨} &(¬ φ₁ ∨ φ₂) &=
        \begin{cases}
         ⊤, &\text{ if } φ₁ ∈_{∨} φ₂\\
         ⊤, &\text{ if } \rm{ndisj}_{∨}~φ₂≡ ⊤\\
         ¬ φ₁ ∨ \rm{ndisj}_{∨}~φ₂, &\text{ otherwise.}
        \end{cases}\\

  \rm{ndisj}_{∨} &(φ₁ ∨ φ₂)&=
        \begin{cases}
         ⊤, &\,\,\,\,\,\,\text{ if } ¬φ₁ ∈_{∨} φ₂ \\
         ⊤, &\,\,\,\,\,\,\text{ if } \rm{ndisj}_{∨}(φ₂) ≡ ⊤\\
         φ₁ ∨ \rm{ndisj}_{∨}~φ₂, &\,\,\,\,\,\,\text{ otherwise.}
        \end{cases}\\
    \rm{ndisj}_{∨}&φ &=φ
    \end{array}
  \end{aligned}
\end{equation}

@@

\begin{mainlemma}[lem-build$_{\boldsymbol{∨}}$]
\label{lem:build-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{build}_{∨}~φ~ψ$.
\end{mainlemma}

@@

\begin{mainlemma}[lem-factor]
\label{lem:factor}
  $Γ ⊢ φ$ then $Γ ⊢ \ffactor(φ)$.
\end{mainlemma}

@@

\begin{mainlemma}
\label{lem:sbuild-or}
If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{sbuild}_{∨}(φ, ψ)$.
\end{mainlemma}

@@

\begin{mainlemma}
  \label{lem:reorder-or}
  If $Γ ⊢ φ$ and $ψ: \Prop$ then $Γ ⊢ \freorder_{∨}~φ~ψ~$, where,
  \begin{align*}
    \label{eq:reorder-or}
    \begin{split}
    &\freorder_{∨} : \Prop \to \Prop \to \Prop\\
    &\freorder_{∨}~φ~ψ~= \rm{sbuild}_{∨}~(rassoc_{∨}~φ)~ψ.
    \end{split}
  \end{align*}
\end{mainlemma}

\begin{proof}
Use Lemma~\ref{lem:rassoc} and Lemma~\ref{lem:sbuild-or}.
\end{proof}

@@

\begin{mainlemma}
  \label{lem:reorder-and}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧}(φ, ψ)$.
\end{mainlemma}

@@

\begin{mainlemma}
  \label{lem:conjunct-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fconjunct_{∨}(φ, ψ)$.
\end{mainlemma}

\begin{mainlemma}
  \label{lem:reorder-and-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧∨}(φ, ψ)$, where
   \begin{align*}
      \begin{split}
      &\freorder : \Prop \to \Prop \to \Prop\\
      &\freorder_{∧∨}~φ~ψ~=
        \begin{cases}
          ψ, &\text{ if } φ≡ψ\\
          ψ, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂,\ ψ₁ ≡ \freorder_{∧∨}~φ~ψ₁~\\
             &\text{and } ψ₂ ≡ \freorder_{∧∨}~φ~ψ₂~\\
          φ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂\\
          \fconjunct_{∨}~φ~ψ~ &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{mainlemma}

@@

To emulate this rule, we perform a sequence of rearrangements inside the
formulas involved to match with the pattern expected by the resolution theorem
described above. We begin by defining a convenient version of
the binary resolution in Def.~\ref{eq:rsol}.

% \begin{equation}
% \label{eq:rsol}
% \begin{aligned}
% &\rm{rsol} : \Prop \to \Prop\\
%       \begin{split}
%         &\rm{rsol} : \Prop \to \Prop\\
%         &\rm{rsol}~φ =
%         \begin{cases}
%           φ₂, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁, φ₂ ≡ φ₄\\
%           φ₂ ∨ φ₄, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∨ (φ₃ ∨ φ₄), φ₃ ≡ ¬ φ₁\\
%           φ, &\text{ otherwise.}
%         \end{cases}
%       \end{split}
%   \end{align*}
% \end{definition}

\begin{mainlemma}
  \label{lem:rsol}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rsol}~φ$.
\end{mainlemma}

@@

\begin{mainth}
  \label{thm:resolve}
  If $Γ ⊢ φ₁$, $Γ ⊢ φ₂$, a literal $l : \Prop$, and $ψ : \Prop$ then
  $Γ ⊢ \fresolve~φ₁~φ₂~l~ψ$, where,
  \begin{equation}
  \begin{split}
  \label{eq:resolve}
    &\fresolve : \Prop \to \Prop \to \Prop \to \Prop \to \Prop\\
    &\fresolve~φ₁~φ₂~l~ψ =
      \rm{rsol}~(\freorder_{∨}~φ₁~(l ∨ ψ) ∧ \freorder_{∨}~φ₂~(¬ l ∨ ψ)).
  \end{split}
  \end{equation}
\end{mainth}

\Needspace{3\baselineskip}
\begin{proof}
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ₁$}
    \RightLabel{Lemma~\ref{lem:reorder-and-or}}
    \UnaryInfC{$Γ ⊢ \freorder_{∨}~φ₁~(l ∨ ψ)$}
    \RightLabel{Lemma~\ref{lem:reorder-and-or}}
    \AxiomC{$Γ ⊢ φ₂$}
    \UnaryInfC{$Γ ⊢ \freorder_{∨}~φ₂~(¬ l ∨ ψ)$}
    \RightLabel{∧-intro}
    \BinaryInfC{$Γ ⊢ \freorder_{∨}~φ₁~(l ∨ ψ) ∧ \freorder_{∨}~φ₂~(¬ l ∨ ψ)$}
    \RightLabel{Lemma~\ref{lem:rsol}}
    \UnaryInfC{$Γ ⊢ \rm{rsol}~(\freorder_{∨}~φ₁~(l ∨ ψ) ∧
     \freorder_{∨}~φ₂~(¬ l ∨ ψ))$}
    \RightLabel{by (\ref{eq:resolve})}
    \UnaryInfC{$Γ ⊢ \fresolve~φ₁~φ₂~l~ψ$}
  \end{bprooftree}
  \end{equation*}
\end{proof}

\begin{example}COMPLETAR ESTE EJEMPLO AL FINAL
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ ¬ r ∨ p ∨ q$}
  \AxiomC{$Γ ⊢ p ∨ q ∨ r $}
  \RightLabel{Theorem \ref{thm:resolve}}
  \BinaryInfC{$Γ ⊢ \fresolve~(¬ r ∨ p ∨ q)~(p ∨ q ∨ r)~r~(p ∨ q)$}
  \UnaryInfC{$Γ ⊢ p ∨ q$}
  \end{bprooftree}
\end{equation*}
\end{example}
